Prevenção Automática de Ataques de Não-Terminação

### 2013-04-24 doc/Manuscript/SBLP13/paper.tex
Including numbers in table 1 and 2

\usepackage[applemac]{inputenc}
\usepackage[brazil]{babel}

@@

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
We say that a program is vulnerable to a non-termination attack if (i) it
contains a loop that is bounded by values coming from public inputs, and (ii)
an adversary can manipulate these values to force this loop to iterate forever.
Preventing this kind of attack is difficult because they do not originate
from bugs that break the semantics of the programming
language, such as buffer overflows.
Instead, they usually are made possible by the wrapping integer arithmetics
used by C, C++ and Java-like languages, which have well-defined semantics.
In this paper we present the diagnosis and the cure for this type of attack.
Firstly, we describe a tainted-flow analysis that detects non-termination
vulnerabilities.
Secondly, we provide a compiler transformation that inserts arithmetic guards on
loop conditions that may not terminate due to integer overflows.
We have implemented our framework in the LLVM compiler, and have tested it on
a benchmark suite containing over 2.5 million lines of C code.
We have found out that the typical path from inputs to loop conditions is,
on the average, XX instructions long.
Our instrumentation that prevents this kind of attacks adds less than XX\%
extra code on the secured program.
The final, protected code, is, on the average, less than XX\% slower than the
original, unprotected program.
\end{abstract}

@@

\begin{figure}[t!]
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|} \hline
Benchmark & Num. Laços & Alcançáveis & Vulneráveis &
Caminhos \\ \hline
470.lbm        & & & &  \\ \hline
433.milc       & & & &  \\ \hline
444.namd       & & & &  \\ \hline
447.dealII     & & & &  \\ \hline
450.soplex     & & & &  \\ \hline
464.h264ref    & & & &  \\ \hline
473.astar      & & & &  \\ \hline
458.sjeng      & & & &  \\ \hline
429.mcf        & & & &  \\ \hline
471.omnetpp    & & & &  \\ \hline
403.gcc        & & & &  \\ \hline
445.gobmk      & & & &  \\ \hline
462.libquantum & & & &  \\ \hline
401.bzip2      & & & &  \\ \hline
456.hmmer      & & & &  \\ \hline
\end{tabular*}
\end{center}
\caption{Informações estáticas inferidas pela análise de não-terminação.
Num. laços: número de laços no programa.
Alcançáveis: quantidade de laços que são dependentes de dados produzidos a
partir de canais de entrada.
Vulneráveis: número de laços que preenchem nossos requisitos de
vulnerabilidade.
Caminhos: tamanho médio do menor caminho de dependência de dados da entrada
até a operação de controle do laço.}
\label{fig:staticLoopData}
\end{figure}

\begin{figure}[t!]
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|} \hline
Benchmark & Num. Instruções & Num. Arit. & Instrumentação &
Crescimento \\ \hline
470.lbm        & & & &  \\ \hline
433.milc       & & & &  \\ \hline
444.namd       & & & &  \\ \hline
447.dealII     & & & &  \\ \hline
450.soplex     & & & &  \\ \hline
464.h264ref    & & & &  \\ \hline
473.astar      & & & &  \\ \hline
458.sjeng      & & & &  \\ \hline
429.mcf        & & & &  \\ \hline
471.omnetpp    & & & &  \\ \hline
403.gcc        & & & &  \\ \hline
445.gobmk      & & & &  \\ \hline
462.libquantum & & & &  \\ \hline
401.bzip2      & & & &  \\ \hline
456.hmmer      & & & &  \\ \hline
\end{tabular*}
\end{center}
\caption{Impacto da instrumentação no código saneado.
Num. Instruções: número de instruções no benchmark.
Num. Arit: número de instruções que podem causar estouro de precisão inteira.
Instrumentação: quantidade de testes inseridos para sanear o programa.
Crescimento: razão entre o tamanho do programa instrumentado e o tamanho do
programa original.}
\label{fig:instImpact}
\end{figure}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[brazil]{babel}

@@

%mmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmmm
\begin{abstract}
We say that a program is vulnerable to a non-termination attack if (i) it
contains a loop that is bounded by values coming from public inputs, and (ii)
an adversary can manipulate these values to force this loop to iterate forever.
Preventing this kind of attack is difficult because they do not originate
from bugs that break the semantics of the programming
language, such as buffer overflows.
Instead, they usually are made possible by the wrapping integer arithmetics
used by C, C++ and Java-like languages, which have well-defined semantics.
In this paper we present the diagnosis and the cure for this type of attack.
Firstly, we describe a tainted-flow analysis that detects non-termination
vulnerabilities.
Secondly, we provide a compiler transformation that inserts arithmetic guards on
loop conditions that may not terminate due to integer overflows.
We have implemented our framework in the LLVM compiler, and have tested it on
a benchmark suite containing over 2.5 million lines of C code.
We have found out that the typical path from inputs to loop conditions is,
on the average, less than 10 instructions long.
Our instrumentation that prevents this kind of attacks adds less than 5\%
extra code on the secured program.
The final, protected code, is, on the average, less than 2\% slower than the
original, unprotected program.
\end{abstract}

@@

\begin{figure}[t!]
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|} \hline
Benchmark & Num. Laços & Alcançáveis & Vulneráveis &
Caminhos \\ \hline
433.milc	&	326	&	8	&	4	&	3   \\ \hline
444.namd	&	508	&	7	&	1	&	2  \\ \hline
447.dealII	&	4657	&	540	&	225	&	3  \\ \hline
450.soplex	&	547	&	21	&	12	&	3  \\ \hline
470.lbm		&	17	&	0	&	0	&	- \\ \hline
401.bzip2	&	205	&	1	&	1	&	3  \\ \hline
403.gcc		&	3853	&	109	&	59	&	4  \\ \hline
429.mcf		&	38	&	2	&	0	&	2  \\ \hline
445.gobmk	&	1073	&	37	&	19	&	4  \\ \hline
456.hmmer	&	687	&	133	&	48	&	2  \\ \hline
458.sjeng	&	235	&	9	&	3	&	4  \\ \hline
462.libquantum	&	67	&	2	&	1	&	4  \\ \hline
464.h264ref	&	1586	&	27	&	22	&	15  \\ \hline
473.astar	&	86	&	7	&	7	&	18  \\ \hline
\end{tabular*}
\end{center}
\caption{Informações estáticas inferidas pela análise de não-terminação.
Num. laços: número de laços no programa.
Alcançáveis: quantidade de laços que são dependentes de dados produzidos a
partir de canais de entrada.
Vulneráveis: número de laços que preenchem nossos requisitos de
vulnerabilidade.
Caminhos: tamanho médio do menor caminho de dependência de dados da entrada
até a operação de controle do laço.}
\label{fig:staticLoopData}
\end{figure}

\begin{figure}[t!]
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|} \hline
Benchmark & Num. Instruções & Num. Arit. & Instrumentação &
Crescimento \\ \hline
433.milc	&	22615	&	1165	&	7	&	0,50\%   \\ \hline
444.namd	&	65045	&	3136	&	1	&	0,03\%   \\ \hline
447.dealII	&	415756	&	14948	&	279	&	0,58\%   \\ \hline
450.soplex	&	60772	&	1815	&	19	&	0,46\%   \\ \hline
470.lbm	&	3614	&	1130	&	0	&	0,00\%   \\ \hline
401.bzip2	&	16359	&	1688	&	1	&	0,10\%   \\ \hline
403.gcc	&	792039	&	14853	&	92	&	0,14\%   \\ \hline
429.mcf	&	2541	&	158	&	0	&	0,00\%   \\ \hline
445.gobmk	&	130183	&	12158	&	54	&	0,63\%   \\ \hline
456.hmmer	&	61908	&	3585	&	86	&	2,00\%   \\ \hline
458.sjeng	&	27708	&	2438	&	20	&	1,10\%   \\ \hline
462.libquantum	&	5848	&	593	&	6	&	1,04\%   \\ \hline
464.h264ref	&	131905	&	13413	&	248	&	2,28\%   \\ \hline
473.astar	&	8261	&	639	&	9	&	1,73\%   \\ \hline
\end{tabular*}
\end{center}
\caption{Impacto da instrumentação no código saneado.
Num. Instruções: número de instruções no benchmark.
Num. Arit: número de instruções que podem causar estouro de precisão inteira.
Instrumentação: quantidade de testes inseridos para sanear o programa.
Crescimento: razão entre o tamanho do programa instrumentado e o tamanho do
programa original.}
\label{fig:instImpact}
\end{figure}
